### 2006-01-22 article/Article.tex
add images to the article
darcs-hash:20060122210145-2fc9d-4b5b4cc3fe3c5741808d7e2eb9d7566e31773845.gz

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}

%  This is a template for Papers to the Insight Journal. 
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{}
\authoraddress{}

\begin{document}
\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

\begin{abstract}
\noindent
% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\end{abstract}

\tableofcontents

% \url{http://www.itk.org}
% \code{Insight/Documentation/Style.pdf}
% \section{Principles of Solar Spot Detection}
% \cite{ITKSoftwareGuide}.
% \doxygen{ImageToImageFilter}

% \small \begin{verbatim}
% \end{verbatim} \normalsize

% The {itemize} environment uses a bullet for each \item.  If you want the 
% \item's numbered, use the {enumerate} environment instead.
% \begin{itemize}
%   \item  Insight Toolkit 2.4.
%   \item  CMake 2.2
% \end{itemize}

% \ref{cthead1}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum0}
\caption{Maximum projection, axe=0.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum1}
\caption{Maximum projection, axe=1.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum}
\caption{Maximum projection, axe=2.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{minimum}
\caption{Minimum projection}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{mean}
\caption{Mean projection.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{median}
\caption{Median projection}
\end{figure}

% \begin{figure}[htbp]
% \centering
% \includegraphics[scale=0.75]{sum}
% \caption{Sum projection}
% \end{figure}

\appendix

\bibliographystyle{plain}
\bibliography{InsightJournal}
\nocite{ITKSoftwareGuide}

\end{document}